---
title: "Faezeh Report"
format: html
editor: visual
---

#### Load data

```{r}
library(readxl)
data <- read_excel("dataset01.xlsx")
test <- read_excel("test.xlsx")

data$STARTING_PRICE <- data$STARTING_PRICE / 1000000
data[c("REGION", "TYPE", "BALCONY")] <- lapply(data[c("REGION", "TYPE", "BALCONY")], as.factor)
levels(data$BALCONY) <- c("0", "1")
```

# Task 3

***Estimate the proportion of apartments in the region through a 90% confidence interval.***

#### Data and problem description

To estimate the proportion of apartments, the following data is used:

-   **Sample Size (n)**: The total number of housing units in the dataset.

-   **Successes**: The number of housing units categorized as "Apartment".

The formula for the 90% confidence interval for a proportion is:

$$ \hat{p} \pm z_{0.05} \sqrt{\frac{\hat{p}(1 - \hat{p})}{n}} $$

Where:

-   $\hat{p}$ is the sample proportion (calculated as the number of successes divided by the sample size).
-   $z_{0.05}$ is the critical value for a 90% confidence level (z = 1.645)).
-   $n$ is the sample size.

```{r}
phat  <- sum(data$TYPE == "Apartment") / n
z     <- qnorm(0.95) # Two-sided, so 90% ci uses z = 0.95
se    <- sqrt(phat*(1-phat)/n)
```

#### Using the formula

The confidence interval is calculated as:

```{r}
conf_int <- c((phat - z*se), phat + z*se)
print(conf_int)
```

#### Using the built-in functions

To validate the manually calculated confidence interval, R's `prop.test()` function is used:

```{r}
ci <- prop.test(sum(data$TYPE == "Apartment"), n, conf.level = 0.9)
ci$conf.int
```

#### Comparing both methods

The results of the manual calculation and the R function are compared to ensure consistency:

```{r}
print(paste0("C.I. proportion of apartments with R: [",round(ci$conf.int[1],3), ", ",round(ci$conf.int[2],3),"]"))
print(paste0("C.I. proportion of apartments by hand: [",round(conf_int[1],3), ", ",round(conf_int[2],3),"]"))
```

#### Conclusion

The 90% confidence interval for the proportion of housing units that are apartments provides a range within which the true proportion is expected to fall with 90% certainty. The manually calculated confidence interval and the interval obtained using R both indicate that the proportion of apartments is approximately 76% to 82%. This insight helps the agency understand how common apartment-style housing is in the area.

# Task 4

***Estimate the expected number of rooms of the housing units in the region through a 95% confidence interval.***

#### Data and problem description

To estimate the expected number of rooms, the following data is used:

-   **Sample Mean (**$\bar{x}$**)**: The average number of rooms across all housing units.

-   **Sample Standard Deviation (s)**: The standard deviation of the number of rooms.

-   **Sample Size (n)**: The total number of housing units.

Since the population variance is unknown, the distribution of the expected number of rooms follows a t-distribution. The formula for the 95% confidence interval is: $$ \bar{x} \pm t_{0.025, n-1} \frac{s}{\sqrt{n}} $$ Where:

-   $t\_{0.025, n-1}$ is the critical value of the t-distribution with (n-1) degrees of freedom at a 95% confidence level.
-   $\bar{x}$ is the sample mean. - (s) is the sample standard deviation.
-   $n$ is the sample size.

```{r}
xbar <- mean(data$ROOMS)  # About 3.15
se   <- sd(data$ROOMS)    # About 1.59
t975 <- qt(0.975, n-1)
```

#### Using the formulas

The confidence interval is calculated as:

```{r}
conf_int <- c((xbar - t975 * se / sqrt(n)), xbar + t975 * se / sqrt(n))
print(conf_int)
```

#### Using the built-in functions

To validate the manually calculated confidence interval, R's `t.test()` function is used:

```{r}
ci <- t.test(data$ROOMS, conf.level = 0.95)
ci$conf.int
```

#### Conclusion

The confidence interval for the mean (expected) number of rooms provides a range within which the true mean is expected to fall with 95% certainty. Both the manually calculated confidence interval and the interval obtained using R suggest that the expected number of rooms is between approximately 3.03 and 3.27. This insight helps the agency understand typical housing configurations in the region.
